%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%\documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% The graphicx package provides the includegraphics command.
\usepackage{graphicx}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
%% following packages allows to write algorithms as pseudocode
\usepackage{amsmath}
\usepackage{algorithmicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\algnewcommand\algorithmicinput{\textbf{INPUT:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\algorithmicoutput{\textbf{OUTPUT:}}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage{lineno}

% math tools (ceiling signs, see: http://tex.stackexchange.com/questions/42271/floor-and-ceiling-functions)
\usepackage{mathtools}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% new commands
\newcommand{\mtrx}[1]{\mathbf{#1}}

\journal{Advances in Engineering Software}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

\title{Singular Value Decomposition Used for Compression of Results from the Finite Element Method}

\author{\v{S}t\v{e}p\'{a}n Bene\v{s}} % Stepan Benes
\ead{stepan.benes@fsv.cvut.cz}
\author{Jaroslav Kruis}
%\author{Jaroslav Kruis\corref{correspondingauthor}}
%\cortext[correspondingauthor]{Corresponding author. Tel.: + 420 224354580.}
\ead{jaroslav.kruis@fsv.cvut.cz}

\address{Department of Mechanics, Faculty of Civil Engineering, Czech Technical University in Prague Th\'akurova 7, Prague 166 29, Czech Republic}

\begin{abstract}
%% Text of abstract should have three parts: Motivation. This paper. Summary.
A complex finite element analysis can produce large amount of data that is problematic to post-process in reasonable time. This paper describes application of Singular Value Decomposition (SVD) to the compression of results from finite element solvers. Although the idea of image compression method is an inspiration for this research work, the SVD compression algorithm used for compression of images cannot be directly used for FEM results. Differences and implementation challenges are discussed in the text. Quality of approximation is more important in scientific field then in computer graphics where the most significant factor is the human perception of the resulting image. Error estimation methods used during compression of finite element results are presented. The focus is also on the algorithm performance. SVD is a very computational intensive method. Therefore, various optimization techniques were investigated, e.g. randomized SVD. The method leads to the lower memory consumption, 10\% of the original size or less, with negligible compression error.
\end{abstract}

\begin{keyword}
Finite Element Method (FEM) \sep Compression \sep Singular Value Decomposition (SVD) \sep Randomized SVD \sep Data Visualization \sep Data Post-processing
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
%\linenumbers

%% main text

% State of the art in FEM data compression
\input{sections/introduction.tex}

% Mathematical background -- SVD, Low-rank approximation matrix, Randomized SVD
\input{sections/math.tex}

% Description of FEM data compression algorithm
\input{sections/implementation.tex}

% Results of compression method (images from postprocessor, tables with compression ratio and error, etc.)
\input{sections/results.tex}

% Evaluation of results
\input{sections/conclusion.tex}

% Acknowledgement
\input{sections/acknowledgement.tex}

%% References with bibTeX database
\section*{References}

\bibliographystyle{model1-num-names}
\bibliography{references}

\end{document}
